---
title: "miRNetR: A companion R package for the miRNet web server"
author: "Le Chang and Jianguo Xia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{miRNetR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 1.0 Overview of miRNetR

miRNetR is an R package, synchronized with the popular miRNet web server, designed for microRNA(miRNA)-centric network analytics and systems-level interpretation. This R package contains the numerous R functions and libraries underlying the web server necessary to perform miRNA data processing and analysis. This package provides support to map from miRNAs to targets (forward mapping), targets to the associated miRNAs (reverse mapping), as well as to generate network files and perform functional enrichment analysis.  

miRNetR is comprised of three modules: (i) Mapping (from miRNAs to targets or from targets to the associated miRNAs), (ii) miRNA network generation, and (iii) Functional enrichment analysis. 

Following installation and loading of miRNetR, users will be able to reproduce web server results from their local computers using the corresponding R command history downloaded from miRNet, thereby achieving maximum flexibility and reproducibility.

## 1.1 Loading the package

After following the installation instructions on the miRNetR Github, you will be ready to use the package. Use the library() function to load the package into R. 

```{r, eval=FALSE}
# Load miRNetR
library(miRNetR)
```

## 1.2 Tips for using the miRNetR package

1) The first function that you will use in every module is the *Init.Data* function, which constructs the dataSetObj object that stores user's data for further processing and analysis. On the miRNet web-server, the suggested name of the dataSetObj, which must be called consistently in your workflow, is *dataSet*. It is not necessary to use this format. Users can use any name they prefer, as long as it is called exactly the same in each step.  

2) The miRNetR package directly creates data files/tables/analysis/networks outputs in your current working directory. 

3) Every command must be run in sequence, please do not skip any commands as this will result in errors downstream. 

4) Each main function in miRNetR is documented. Use the ?Function format to open its documentation. For instance, use ?miRNetR::PerformMirGeneMapping to find out more about this function.

## 1.3 Data Format

miRNetR supports the upload of miRNA or targets list data. These files can be uploaded in a simple text format (.txt or .csv format). 

## 1.4 Reading in Data Files

The first step of any analysis is to read in the data. miRNetR provides multiple example datasets for testing purposes. For this tutorial, we will use a list of miRNAs as input data. 

```{r, eval=FALSE}
# Initiate the dataSetObj
dataSet<-Init.Data("mir", "mirlist");

# Set up the miRNA list data
dataSet<-SetupMirListData(mirs="hsa-mir-101-3p
hsa-mir-133b
hsa-mir-147a
hsa-mir-3140-3p
hsa-mir-361-5p
hsa-mir-510-5p", orgType="hsa", idType="mir_id", tissue="Kidney");
```

## 2.1 Map miRNAs to Targets

```{r, eval=FALSE}
# Map from miRNAs to their target genes
dataSet<-QueryMultiListMir();
```

## 2.2 Network Generation

```{r, eval=FALSE}
# 
dataSet<-CreateMirNets("multilist");
PrepareMirNet("mirnet1", "mirnet_0.json");
```

## 3. Functional Enrichment Analysis

```{r, eval=FALSE}
#  
PerformMirTargetEnrichAnalysis(adjust.type="NA", fun.type="mirfamily", file.nm="network_enrichment_mirfamily_1", IDs="hsa-mir-101-3p; hsa-mir-147a; hsa-mir-361-5p; hsa-mir-133b; hsa-mir-510-5p; hsa-mir-3140-3p", algo="hyp")
```
